Over the course of my graduate studies, and even before that as a non-degree seeking student, I've picked up as many mentors as I possibly could. My list is long and I'm thrilled to now have the opportunity to recognize every person and relive those experiences.

Thank you to my supervisors, Katrin Erk and Jessy Li. Katrin, thank you for introducing me to Computational Linguistics and then taking the leap to be my supervisor. Thank you for giving me free reigns over my research while also ensuring I was financially supported with RAships and helping me shape my successful NSF application. You helped me strike a balance between strongly motivated and developed research (coming back to the "why?"), and a commitment to taking time off (turning into pumpkins at 10pm). Jessy, thank you for a deluge of computational discourse knowledge and I am proud to be your first Linguistics PhD advisee. Thank you to my committee members, Greg Durrett, Amanda Stent and Hans Kamp. Greg, thank you for being my co-co-supervisor and helping me in every aspect from technical coding problems to tips on presenting at conferences. Amanda, thank you for your probing feedback even during my proposal, which prompted me to take a closer look at speech acts and reevaluate my initial framing of subjective annotations. Hans, thank you for stepping outside of your research and asking insightful questions that led to interesting extensions of my research.

Many other faculty outside of my committee played a pivotal role in my studies: Rajka Smiljanic, Colin Bannard, and Byron Wallace. Rajka, thank you for taking me under your wing and giving me the first taste of computational approaches in Linguistics, sparking an interest in statistics, experimental design and spectrograms. I wish I could have gotten a second PhD in this area. Thank you, Colin, for helping me embark on my first annotation project and my first experience with Wikipedia (perhaps both obligatory for computational linguists?). Thank you, Byron, for helping me carry out and publish my first research in Computational Linguistics, and for opening my eyes to the field of biomedical NLP with both practical and beneficial applications.

My fellow graduate students, both senior and junior, provided an invaluable source of knowledge and emotional support: Eric Holgate, Su Wang, Alex Rosenfeld, Pengxiang Chang, Stephen Roller. Thank you, Eric, for being my commiserating buddy and for insightful conversations on entities. Su, thanks for your dedication to turn a class project into a paper. Alex, thank you for always offering to help and always offering your opinion. Thank you, Pengxiang, for helping me remember how to compile C and answering all my TACC/Chameleon questions. Thank you, Stephen, for your help with the CS machines and for proofreading our paper.

Finally, I recognize my deepest sources of inspiration and support: my husband, my son, and my family. Thank you, Noel, for encouraging me to write down and brainstorm what started it all: my crazy 10-year vision of becoming a research scientist who is featured on NPR. Thank you to my little cucumber who is not so little anymore and who has helped me achieve a balanced graduate life, letting me step into your world of make-believe on our monster walks and countless invented games.


%\begin{flushright}
%~\\
%Stephen Roller\\
%Austin, Texas\\
%April 2017
%\end{flushright}
